The transboundary nature of the world’s exploited marine species

### 2020-11-25 Scripts/Manuscript_Draft.Rmd
we for I

Forty years after the formal adoption of UNCLOS [@UnitedNations:1986tl] and the subsequent 1995 United Nations Fish Stock Agreement for the cooperation on the management and conservation of straddling and highly migratory resources [@UnitedNations:1995fs], accurate estimates of the number of exploited marine species shared by neighboring nations are still unavailable. An informed guess, based on limited biogeographical data, suggested that there are approximately 500 to 1500 exploited transboundary stocks [@Caddy:1997ue]. A recent literature review included 344 shared taxa and their historical contributions to fisheries [@Teh:2015gd]. However, these studies did not consider species' biogeography in the quantification of transboundary stocks. Here, we aim to estimate the number of exploited marine species shared by neighboring countries and determine their contribution to global and regional catches as well as fisheries revenue. Moreover, we categorize species according to their catch trends while identifying differences among species based on habitat preference. We hypothesize that the methodological constraints of previous studies resulted in an underestimation of the number of transboundary species and their contribution to global catch and revenue.

We overlaid the known distribution of 938 commercially valuable marine species responsible for an average of 96.5% of global EEZ catches between 2005 and 2014, and 280 EEZs of 198 coastal countries (See Methods). We define a ‘stock’ unit as a species in an EEZ, instead of a genetically or morphological distinct unit [@Nelson:2016kk], due to the lack of such biological information being available for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level (or even as groups containing multiple species). For example, shrimp (e.g., *Litopenaeus stylirostris* or *Farfantepenaeus californiensis*) along the Pacific coast of Mexico [@DiarioOficialdelaFederacionDOF:2018wt] and hammerhead shark (*Sphyrna zygaena*) in Peru [@MAP:2017uf] are managed at the species level, yet include multiple populations exploited by each country’s fisheries. Moreover, recent research shows connectivity across fish stocks through larval dispersal [@Ramesh:2019va] and adult migration [@Levin:2018in; @Popova:2019fs; @Dunn:2019ca], although considerable level of uncertainty exists at different life stages [@Kaplan:2016hs; @Archambault:2016ie]. For this analysis we only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g., we did not include the high seas). We rely on multiple data sources including occurrence, distribution models and catch data, and only consider a species to be present in a grid cell if all data sources showed positive occurrence (see Methods).

@@

The Sea Around Us has reconstructed global fisheries catches from 1950 to 2014, identifying commercial marine fish and invertebrates as well as fishing regions [@Zeller2016]. We used the Sea Around Us reconstructed species list to determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs. The total number of species analyzed was 938, representing 67% of identified species by the Sea Around Us and accounting for 96.52% of the catch identified at the species’ level. To determine the current distribution of exploitable marine species, we used four data sources of species-distributions: (i) occurrence data, (ii) an Ensemble Environmental Niche Model (ENMs), (iii) a life-history-based distribution model, and (iv) fisheries catch data (Table \@ref(tab:ffvTable1)). Each source represents a different method of estimating the distribution of a given species, thus providing a more robust result than an analysis focused on a single data source. Only commercially fished species with data from all four sources were included in our analysis.

@@

In addition to occurrence data, we used two different methods to estimate species distributions, hereafter referred to as Ecological Niche Model-Nereus (ENM-Nereus) and Species Distribution Model-SAU (SDM-SAU). Although they use the same occurrence and environmental data, the models are structurally different complementing each other and providing robustness to the results.

@@

The previous models combine observational data with a series of biotic and abiotic data to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g., ENM-Nereus uses four different species distribution algorithms and SDM-SAU undertakes validation by means of other models), we used a fourth data set to corroborate the models’ outputs. The Sea Around Us estimates total reconstructed catches - catches based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in available FAO data - for each country. Catches are also spatially allocated on a $0.5^{\circ}$ x $0.5^{\circ}$ latitude longitude grid [@Zeller2016]. Roughly, the Sea Around Us method consist of the following steps. First, it takes each country's officially reported catch data (e.g., National, FAO or RFMO). Secondly, it uses literature (e.g., peer review, grey literature) to identify missing components (e.g., species, gears) and sources of alternative information for missing components. It then derives country estimates for missing data and creates time series interpolation. Finally, the estimated and official data are aggregated, making up the total reconstructed catch data (see Pauly *et al* [-@Pauly:2019bf] and Zeller *et al.* [-@Zeller2016] for catch reconstruction and spatial allocation details). We used the Sea Around Us catch reconstruction database from 2005 to 2014 as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs. We selected this time frame to investigate the recent (our time frame includes the last decade of available data) contribution of transboundary fisheries to catches and revenue from fisheries and to reduce the uncertainty embedded in the reconstruction process (see Key Uncertainties). Note that in call cases we report the average catch from 2005 to 2014.

@@

We developed a three-criteria methodology to determine whether or not a species can be considered transboundary. Only species that met all criteria at least once were considered as “transboundary”, while species that did not meet the criteria for any EEZ analyzed were considered as “discrete”. Note that in cases where a species met all criteria for some EEZs, but not for other EEZs, these species were still considered as “transboundary”.

@@

As mentioned above, we define transboundary species as those marine species that occur within the EEZs of two or more neighboring countries. Hence, according to this criteria a species was only considered as transboundary if it was shared between two neighboring countries, regardless of the species extended distribution. The analysis was undertaken only within the boundaries of the EEZs of coastal states using the Sea Around Us shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) - noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico's EEZ is divided in Mexico Pacific and Mexico Atlantic), including islands territories - and determined the intersections between polygons using the R package sf [@PackagesfSimpleF:2018vp]. When estimating transboundary species, we filtered out those shared by EEZs sub-regions (e.g., USA Gulf of Mexico and USA Atlantic), and when aggregating results by country, species that occurred in more than one sub-region were only accounted for once. Species that were present in EEZs that were non-continental territories neighboring other countries were kept as separate (e.g., Argentina and Falkland Islands), but removed in cases where the non-continental territory belonged to the same nation (e.g., Brazil and Fernando de Noronha). Associated states like Puerto Rico and New Caledonia were not considered separately (e.g., Puerto Rico was grouped with the United States and New Caledonia with France).

@@

We used the occurrence database, the ENM-Nereus model, and SDM-SAU model to determine the presence of each species within each of the world's $0.5^{\circ}$ x $0.5^{\circ}$ marine grid cells. All analyses only considered cases with agreement across all databases to obtain a more conservative estimate of transboundary species. Moreover, we assumed that a species was only present in a given grid-cell if it was reported in the Sea Around Us database. Therefore, all species that were not reported as caught in any single year between the reference years (2005 to 2014) in a given grid-cell were dropped. This rule assumes that if a commercial species is projected within the EEZ of any fishing country, such a species would have been fished, and thus likely reported at some point over the last decade of data (2005 - 2014), thereby validating the models and selecting "currently" shared species. We acknowledge that this criteria might limit the global distribution of species therefore resulting in a conservative estimate of transboundary species.

@@

Finally, to have a more robust result and not categorize a species as transboundary based on its presence in a single $0.5^{\circ}$ x $0.5^{\circ}$ grid cell within an EEZ, we computed an Area Index. The Area Index represents the proportion of a given species' overarching shared distribution between neighboring EEZs accounted for by each individual EEZ. We classified a species as transboundary if both neighboring EEZs enclosed over 25% of the species joint shared distribution. While a species that has less than the selected threshold is not considered transboundary in this paper, this threshold can be lowered for a more relaxed result or increased for a more conservative estimate (Figure \@ref(fig:ffvFigureS2)).

@@

We estimated the economic contribution in 2019 real USD of transboundary species for each country using global ex-vessel price data [@Tai2017]. The database we draw from includes ex-vessel price derived from multiple sources and a structured interpolation method (e.g., similar countries, species) to fill in data gaps [@Sumaila:2015uc]. The dataset is harmonized with the Sea Around Us catch data to estimate yearly fishing revenue (as ex-vessel price) for all species and EEZs considered in this study (http://www.seaaroundus.org/data/#/feru). We report average fishing revenue derived from fishing activity within global EEZs between 2005 to 2014. We did not include revenue from areas beyond national jurisdiction. We used the monthly average US Consumer Price Index (CPI) according to the U.S. Bureau of Labor Statistic (https://www.bls.gov/cpi/) to standardize the original 2010 real USD value to 2019 real USD.

We used catch data as described above to determine the catch trend of each species within an EEZ. Although this method has previously been used to estimate stock status [@Grainger:1996tk], the categories presented here are intended to represent catch trends, and not fishing status for each species as many environmental and social-economic factors (e.g., temperature, markets, fishing policies, and fishing effort) affect catches [@Pauly:2013gs; @Branch:2008kk]. We only assessed species within each EEZ for which at least ten years of data were available between 1951 and 2014 and with at least five consecutive years of data. Three final categories were drawn up for each species depending on catch volume within each EEZ (e.g., present, maximum, and minimum EEZ's historical catch) and the year (e.g., year of maximum historical catch of the species within that EEZ) (Table \@ref(tab:ffvTable2)) [@Kleisner:2011wn]. Accordingly, Category A represents fisheries that are registering increases in catch (“increasing”); Category B, species that have a constant catch rate (“constant”); and Category C, species that have registered declines in catch over the last 10 years (“decreasing”). Finally, we report the predominant category over the time period considered.

@@

All analyses were run using the statistical software R version 3.5.2 (2018-12-20) with the packages data.table  [@Packagedatatable:2019uh], janitor [@PackagejanitorSim:2018ur], wesanderson [@Packagewesanderson:2018vn], rfishbase [@PackagerfishbaseR:2019th], R.matlab [@PackageRmatlabRe:2018un], sf [@PackagesfSimpleF:2018vp], sp [@PackagespClasses:2019tn], tidiverse [@PackagetidyverseE:2017vq], tidytext [@PackagetidytextTe:2019vk], and zoo [@PackagezooSInfr:2019tm]. All code is available at https://github.com/jepa/FishForVisa. All maps were made with Natural Earth data available at https://www.naturalearthdata.com/. We performed a series of one-way analysis of variance (one-way ANOVA) and Multivariate analysis of variance (MANOVA) to determine statistically significant differences between the means of different groups (e.g., geographical regions, species, catch trends) of transboundary species and their contribution to catch and fishing revenue [@Chambers:1992up; @Krzanowski:1990wm]. We used the standard R functions for the ANOVA and MANOVA after testing for assumptions. In cases where the ANOVA assumptions were not clearly met, we ran the non-parametric version Kruskal-Wallis Rank Sum Test to confirm results [@Hollander:2013wz].

@@

We have identified four key uncertainties in the method utilized that may affect the estimation of transboundary species. Firstly, as we ran the analysis at the species level due to limited spatial-specific data on species sub populations (stocks), we are not able to identify transboundary stocks within EEZs. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level. Also, recent research suggests strong connectivity between stocks, even when separated by thousands of kilometers [@Ramesh:2019va] providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca]. However, it is important to acknowledge that there is considerable uncertainty associated in determining levels of connectivity across time and space for marine populations at different life stages, from larvae [@Kaplan:2016hs] to adults [@Archambault:2016ie]. This is of special concern, but-not-limited-to [@Archambault:2016ie], highly migratory species whose ranges span multiple jurisdictions and the high seas, such as tunas [@Moore:2020cd], billfishes [@Sepulveda:2019hq] and sharks [@Vandeperre:2014jk] challenging management decisions based on meta-populations [@Moore:2020cd]. Here, we limited the definition of “transboundary” to include species spanning only adjacent countries (e.g., the US and Canada), excluding countries that were separated by another nation (e.g., Canada and Mexico) and/or the high seas (e.g., Canada and Spain). Consequently, our results likely provide a conservative estimate of transboundary species, as we did not cover all marine taxa in the world [@Reygondeau:2019uh] and only analyzed species present in all four data sources. Nevertheless, our results are representative of a substantial proportion of the world's marine catches and revenue from economically important marine fisheries. Secondly, the predicted species' distribution is affected by the uncertainties of the environmental data and models used for such predictions. Structural differences within Earth System Models result in variations of oceanic conditions, which undoubtedly affect the ENM-Nereus. We averaged results from the three models in an effort to capture the structural variation across models. Natural climate variability is a major driver of marine species' distributions, potentially removing a species from an EEZ for long periods (e.g., Anchovies and Sardines "regimes" in the Eastern Pacific are strongly influenced by water temperature decal oscillations [@Chavez:2003jr]. Thus, a species’ distribution can potentially be reduced or shifted in such a way that it only covers one EEZ until oceanic conditions change again and the associated species' distribution expands. To account for such climate variability, we derived the ENM-Nereus results as an average of oceanic conditions between 1970 and 2000. This is not an issue for the SDM-SAU as it does not directly require environmental variables [@Pauly:2016uz]. Thirdly, we assumed that, if the Sea Around Us reconstructed data recorded a species as caught in any given grid cell, then the species was present within that grid cell. While catch data are not exempt of uncertainty, in most cases, differences between the Sea Around Us and the FAO self-reported data are smaller towards the end of the time series [@Pauly:2019bf]. Thus, we limited the catch data reference period in our analyses to between 2005 and 2014, the last ten years of data available. Likewise, the spatial allocation of the catch is subjected to imprecision, mainly produced by differences in the spatial scale of the original data and the method employed by the Sea Around Us. Finally, this study’s results only considered species for which all datasets agreed on presence and had and Area Index of at least 25% (e.g., the species shared distribution was at least 25% in each EEZ). Therefore, again, our approach presents a relatively conservative estimate of the number of transboundary species in the world. Using a smaller Area Index will result in more transboundary species (Figure \@ref(fig:ffvFigureS2)).

@@

We identified 633 exploited transboundary species worldwide (67.5% of the 938 species analyzed), almost double previous estimates [@Teh:2015gd]. Between 2005 and 2014, national fleets targeting these transboundary species within EEZs caught an annual average of 48.5 million tonnes, representing 82.3% of EEZ-derived catches reconstructed by the Sea Around Us at the species level (Figure \@ref(fig:ffvFigure1)a). These catches generated a yearly average of USD 77,591 million in fishing revenue (78.5% of global fishing revenue) over the same time period. Our findings are considerably higher than the previous estimates of 34.2 million tonnes in catches and USD 40,187 million (in 2019 value) in fishing revenue from shared stocks [@Teh:2015gd]. When we re-estimated transboundary species’ catches using data consistent with those used in previous studies (FAO global reported data in year 2006), the 633 transboundary species identified here accounted for 40.4 million tonnes of annual catches (i.e., 18% higher than previously estimated). Our results suggest that the contribution of transboundary species to global catch and fishing revenue might previously have been  underestimated due to an incomplete understanding of the transboundary nature of marine fished species. The 305 non-transboundary species (termed as ‘discrete’ species here, see Methods - Determining transboundary species trait) accounted for a much smaller proportion of total catch and revenue; 2.8 million tonnes and USD 4,282 million annually respectively, on average, between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

@@

At a regional level, we found that transboundary species are particularly economically important for Northern America (average country revenue = USD 4,680 $\pm$ 6,000 million) and Eastern Asia (USD 3,779 $\pm$ 3,093 million) (Figure \@ref(fig:ffvFigure1)b). The estimated per country revenues from transboundary species in these two regions is significantly different from other regions (one-way ANOVA; DF(16,165) = 5.081, p < 0.001, $\alpha$ = 0.05). China (USD 7,284 million) and the USA (USD 11,604 million) contribute 55% and 82% to the annual average revenue from 2005 to 2014  in Eastern Asia and Northern America, respectively. In addition to China and the USA, Russia (USD 7,379 million), Peru (USD 6,044 million) and Japan (USD 3,907 million) were among the top five countries with the most fishing revenue generated from transboundary species between 2005 and 2014 (Figure \@ref(fig:ffvFigure1)a). These five nations were responsible for 41% of the yearly global fisheries revenue from transboundary species.

@@

We determined the catch trend of each species within each EEZ, classifying them as increasing (Category A), constant (Category B) or decreasing (Category C) (Figures \@ref(fig:ffvFigure3); \@ref(fig:ffvFigureS2)). While previous studies have demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Our intention here is to examine where the catch trends of transboundary species differ from non-transboundary species between 2005 and 2014 relative to historic catch since 1951 (see Methods). We found significant differences for all catch trend categories for transboundary species (one-way ANOVA, Df(2,459) = 47.94, p < 0.001, $\alpha$ = 0.05; Tukey's post hoc test p $\leq$ 0.001), and no significant differences in catch trend categories for discrete species (one-way ANOVA, Df(2,106) = 1.885, p = 0.157; $\alpha$ = 0.05). We also found significant differences in catch trends when directly comparing transboundary to discrete species categories (MANOVA, Df(2,459) = 19.001, p < 0.001). Overall, transboundary species only targeted by one country are generally less likely to have a decreasing catch trend compared to instances where the shared species is fished by neighboring countries (Table \@ref(tab:ffvTableS2)).

>>>>

Forty years after the formal adoption of UNCLOS [@UnitedNations:1986tl] and the subsequent 1995 United Nations Fish Stock Agreement for the cooperation on the management and conservation of straddling and highly migratory resources [@UnitedNations:1995fs], accurate estimates of the number of exploited marine species shared by neighboring nations are still unavailable. An informed guess, based on limited biogeographical data, suggested that there are approximately 500 to 1500 exploited transboundary stocks [@Caddy:1997ue]. A recent literature review included 344 shared taxa and their historical contributions to fisheries [@Teh:2015gd]. However, these studies did not consider species' biogeography in the quantification of transboundary stocks. Here, I aim to estimate the number of exploited marine species shared by neighboring countries and determine their contribution to global and regional catches as well as fisheries revenue. Moreover, I categorize species according to their catch trends while identifying differences among species based on habitat preference. I hypothesize that the methodological constraints of previous studies resulted in an underestimation of the number of transboundary species and their contribution to global catch and revenue.

I overlaid the known distribution of 938 commercially valuable marine species responsible for an average of 96.5% of global EEZ catches between 2005 and 2014, and 280 EEZs of 198 coastal countries (See Methods). I define a ‘stock’ unit as a species in an EEZ, instead of a genetically or morphological distinct unit [@Nelson:2016kk], due to the lack of such biological information being available for almost all the species included in this study [@Teh:2015gd]. While I acknowledge that species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level (or even as groups containing multiple species). For example, shrimp (e.g., *Litopenaeus stylirostris* or *Farfantepenaeus californiensis*) along the Pacific coast of Mexico [@DiarioOficialdelaFederacionDOF:2018wt] and hammerhead shark (*Sphyrna zygaena*) in Peru [@MAP:2017uf] are managed at the species level, yet include multiple populations exploited by each country’s fisheries. Moreover, recent research shows connectivity across fish stocks through larval dispersal [@Ramesh:2019va] and adult migration [@Levin:2018in; @Popova:2019fs; @Dunn:2019ca], although considerable level of uncertainty exists at different life stages [@Kaplan:2016hs; @Archambault:2016ie]. For this analysis I only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g., I did not include the high seas). I rely on multiple data sources including occurrence, distribution models and catch data, and only consider a species to be present in a grid cell if all data sources showed positive occurrence (see Methods).

@@

The Sea Around Us has reconstructed global fisheries catches from 1950 to 2014, identifying commercial marine fish and invertebrates as well as fishing regions [@Zeller2016]. I used the Sea Around Us reconstructed species list to determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs. The total number of species analyzed was 938, representing 67% of identified species by the Sea Around Us and accounting for 96.52% of the catch identified at the species’ level. To determine the current distribution of exploitable marine species, I used four data sources of species-distributions: (i) occurrence data, (ii) an Ensemble Environmental Niche Model (ENMs), (iii) a life-history-based distribution model, and (iv) fisheries catch data (Table \@ref(tab:ffvTable1)). Each source represents a different method of estimating the distribution of a given species, thus providing a more robust result than an analysis focused on a single data source. Only commercially fished species with data from all four sources were included in our analysis.

@@

In addition to occurrence data, I used two different methods to estimate species distributions, hereafter referred to as Ecological Niche Model-Nereus (ENM-Nereus) and Species Distribution Model-SAU (SDM-SAU). Although they use the same occurrence and environmental data, the models are structurally different complementing each other and providing robustness to the results.

@@

The previous models combine observational data with a series of biotic and abiotic data to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g., ENM-Nereus uses four different species distribution algorithms and SDM-SAU undertakes validation by means of other models), I used a fourth data set to corroborate the models’ outputs. The Sea Around Us estimates total reconstructed catches - catches based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in available FAO data - for each country. Catches are also spatially allocated on a $0.5^{\circ}$ x $0.5^{\circ}$ latitude longitude grid [@Zeller2016]. Roughly, the Sea Around Us method consist of the following steps. First, it takes each country's officially reported catch data (e.g., National, FAO or RFMO). Secondly, it uses literature (e.g., peer review, grey literature) to identify missing components (e.g., species, gears) and sources of alternative information for missing components. It then derives country estimates for missing data and creates time series interpolation. Finally, the estimated and official data are aggregated, making up the total reconstructed catch data (see Pauly *et al* [-@Pauly:2019bf] and Zeller *et al.* [-@Zeller2016] for catch reconstruction and spatial allocation details). I used the Sea Around Us catch reconstruction database from 2005 to 2014 as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs. I selected this time frame to investigate the recent (our time frame includes the last decade of available data) contribution of transboundary fisheries to catches and revenue from fisheries and to reduce the uncertainty embedded in the reconstruction process (see Key Uncertainties). Note that in call cases I report the average catch from 2005 to 2014.

@@

I developed a three-criteria methodology to determine whether or not a species can be considered transboundary. Only species that met all criteria at least once were considered as “transboundary”, while species that did not meet the criteria for any EEZ analyzed were considered as “discrete”. Note that in cases where a species met all criteria for some EEZs, but not for other EEZs, these species were still considered as “transboundary”.

@@

As mentioned above, I define transboundary species as those marine species that occur within the EEZs of two or more neighboring countries. Hence, according to this criteria a species was only considered as transboundary if it was shared between two neighboring countries, regardless of the species extended distribution. The analysis was undertaken only within the boundaries of the EEZs of coastal states using the Sea Around Us shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) - noting that it subdivides the EEZs of 198 coastal states into 280 regions (e.g., Mexico's EEZ is divided in Mexico Pacific and Mexico Atlantic), including islands territories - and determined the intersections between polygons using the R package sf [@PackagesfSimpleF:2018vp]. When estimating transboundary species, I filtered out those shared by EEZs sub-regions (e.g., USA Gulf of Mexico and USA Atlantic), and when aggregating results by country, species that occurred in more than one sub-region were only accounted for once. Species that were present in EEZs that were non-continental territories neighboring other countries were kept as separate (e.g., Argentina and Falkland Islands), but removed in cases where the non-continental territory belonged to the same nation (e.g., Brazil and Fernando de Noronha). Associated states like Puerto Rico and New Caledonia were not considered separately (e.g., Puerto Rico was grouped with the United States and New Caledonia with France).

@@

I used the occurrence database, the ENM-Nereus model, and SDM-SAU model to determine the presence of each species within each of the world's $0.5^{\circ}$ x $0.5^{\circ}$ marine grid cells. All analyses only considered cases with agreement across all databases to obtain a more conservative estimate of transboundary species. Moreover, I assumed that a species was only present in a given grid-cell if it was reported in the Sea Around Us database. Therefore, all species that were not reported as caught in any single year between the reference years (2005 to 2014) in a given grid-cell were dropped. This rule assumes that if a commercial species is projected within the EEZ of any fishing country, such a species would have been fished, and thus likely reported at some point over the last decade of data (2005 - 2014), thereby validating the models and selecting "currently" shared species. I acknowledge that this criteria might limit the global distribution of species therefore resulting in a conservative estimate of transboundary species.

@@

Finally, to have a more robust result and not categorize a species as transboundary based on its presence in a single $0.5^{\circ}$ x $0.5^{\circ}$ grid cell within an EEZ, I computed an Area Index. The Area Index represents the proportion of a given species' overarching shared distribution between neighboring EEZs accounted for by each individual EEZ. I classified a species as transboundary if both neighboring EEZs enclosed over 25% of the species joint shared distribution. While a species that has less than the selected threshold is not considered transboundary in this chapter, this threshold can be lowered for a more relaxed result or increased for a more conservative estimate (Figure \@ref(fig:ffvFigureS2)).

@@

I estimated the economic contribution in 2019 real USD of transboundary species for each country using global ex-vessel price data [@Tai2017]. The database I draw from includes ex-vessel price derived from multiple sources and a structured interpolation method (e.g., similar countries, species) to fill in data gaps [@Sumaila:2015uc]. The dataset is harmonized with the Sea Around Us catch data to estimate yearly fishing revenue (as ex-vessel price) for all species and EEZs considered in this study (http://www.seaaroundus.org/data/#/feru). I report average fishing revenue derived from fishing activity within global EEZs between 2005 to 2014. I did not include revenue from areas beyond national jurisdiction. I used the monthly average US Consumer Price Index (CPI) according to the U.S. Bureau of Labor Statistic (https://www.bls.gov/cpi/) to standardize the original 2010 real USD value to 2019 real USD.

I used catch data as described above to determine the catch trend of each species within an EEZ. Although this method has previously been used to estimate stock status [@Grainger:1996tk], the categories presented here are intended to represent catch trends, and not fishing status for each species as many environmental and social-economic factors (e.g., temperature, markets, fishing policies, and fishing effort) affect catches [@Pauly:2013gs; @Branch:2008kk]. I only assessed species within each EEZ for which at least ten years of data were available between 1951 and 2014 and with at least five consecutive years of data. Three final categories were drawn up for each species depending on catch volume within each EEZ (e.g., present, maximum, and minimum EEZ's historical catch) and the year (e.g., year of maximum historical catch of the species within that EEZ) (Table \@ref(tab:ffvTable2)) [@Kleisner:2011wn]. Accordingly, Category A represents fisheries that are registering increases in catch (“increasing”); Category B, species that have a constant catch rate (“constant”); and Category C, species that have registered declines in catch over the last 10 years (“decreasing”). Finally, I report the predominant category over the time period considered.

@@

All analyses were run using the statistical software R version 3.5.2 (2018-12-20) with the packages data.table  [@Packagedatatable:2019uh], janitor [@PackagejanitorSim:2018ur], wesanderson [@Packagewesanderson:2018vn], rfishbase [@PackagerfishbaseR:2019th], R.matlab [@PackageRmatlabRe:2018un], sf [@PackagesfSimpleF:2018vp], sp [@PackagespClasses:2019tn], tidiverse [@PackagetidyverseE:2017vq], tidytext [@PackagetidytextTe:2019vk], and zoo [@PackagezooSInfr:2019tm]. All code is available at https://github.com/jepa/FishForVisa. All maps were made with Natural Earth data available at https://www.naturalearthdata.com/. I performed a series of one-way analysis of variance (one-way ANOVA) and Multivariate analysis of variance (MANOVA) to determine statistically significant differences between the means of different groups (e.g., geographical regions, species, catch trends) of transboundary species and their contribution to catch and fishing revenue [@Chambers:1992up; @Krzanowski:1990wm]. I used the standard R functions for the ANOVA and MANOVA after testing for assumptions. In cases where the ANOVA assumptions were not clearly met, I ran the non-parametric version Kruskal-Wallis Rank Sum Test to confirm results [@Hollander:2013wz].

@@

I have identified four key uncertainties in the method utilized that may affect the estimation of transboundary species. Firstly, as I ran the analysis at the species level due to limited spatial-specific data on species sub populations (stocks), I are not able to identify transboundary stocks within EEZs. While I acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level. Also, recent research suggests strong connectivity between stocks, even when separated by thousands of kilometers [@Ramesh:2019va] providing additional ecological ground for our analysis [@Popova:2019fs; @Dunn:2019ca]. However, it is important to acknowledge that there is considerable uncertainty associated in determining levels of connectivity across time and space for marine populations at different life stages, from larvae [@Kaplan:2016hs] to adults [@Archambault:2016ie]. This is of special concern, but-not-limited-to [@Archambault:2016ie], highly migratory species whose ranges span multiple jurisdictions and the high seas, such as tunas [@Moore:2020cd], billfishes [@Sepulveda:2019hq] and sharks [@Vandeperre:2014jk] challenging management decisions based on meta-populations [@Moore:2020cd]. Here, I limited the definition of “transboundary” to include species spanning only adjacent countries (e.g., the US and Canada), excluding countries that were separated by another nation (e.g., Canada and Mexico) and/or the high seas (e.g., Canada and Spain). Consequently, our results likely provide a conservative estimate of transboundary species, as I did not cover all marine taxa in the world [@Reygondeau:2019uh] and only analyzed species present in all four data sources. Nevertheless, our results are representative of a substantial proportion of the world's marine catches and revenue from economically important marine fisheries. Secondly, the predicted species' distribution is affected by the uncertainties of the environmental data and models used for such predictions. Structural differences within Earth System Models result in variations of oceanic conditions, which undoubtedly affect the ENM-Nereus. I averaged results from the three models in an effort to capture the structural variation across models. Natural climate variability is a major driver of marine species' distributions, potentially removing a species from an EEZ for long periods (e.g., Anchovies and Sardines "regimes" in the Eastern Pacific are strongly influenced by water temperature decal oscillations [@Chavez:2003jr]. Thus, a species’ distribution can potentially be reduced or shifted in such a way that it only covers one EEZ until oceanic conditions change again and the associated species' distribution expands. To account for such climate variability, I derived the ENM-Nereus results as an average of oceanic conditions between 1970 and 2000. This is not an issue for the SDM-SAU as it does not directly require environmental variables [@Pauly:2016uz]. Thirdly, I assumed that, if the Sea Around Us reconstructed data recorded a species as caught in any given grid cell, then the species was present within that grid cell. While catch data are not exempt of uncertainty, in most cases, differences between the Sea Around Us and the FAO self-reported data are smaller towards the end of the time series [@Pauly:2019bf]. Thus, I limited the catch data reference period in our analyses to between 2005 and 2014, the last ten years of data available. Likewise, the spatial allocation of the catch is subjected to imprecision, mainly produced by differences in the spatial scale of the original data and the method employed by the Sea Around Us. Finally, this study’s results only considered species for which all datasets agreed on presence and had and Area Index of at least 25% (e.g., the species shared distribution was at least 25% in each EEZ). Therefore, again, our approach presents a relatively conservative estimate of the number of transboundary species in the world. Using a smaller Area Index will result in more transboundary species (Figure \@ref(fig:ffvFigureS2)).

@@

I identified 633 exploited transboundary species worldwide (67.5% of the 938 species analyzed), almost double previous estimates [@Teh:2015gd]. Between 2005 and 2014, national fleets targeting these transboundary species within EEZs caught an annual average of 48.5 million tonnes, representing 82.3% of EEZ-derived catches reconstructed by the Sea Around Us at the species level (Figure \@ref(fig:ffvFigure1)a). These catches generated a yearly average of USD 77,591 million in fishing revenue (78.5% of global fishing revenue) over the same time period. Our findings are considerably higher than the previous estimates of 34.2 million tonnes in catches and USD 40,187 million (in 2019 value) in fishing revenue from shared stocks [@Teh:2015gd]. When I re-estimated transboundary species’ catches using data consistent with those used in previous studies (FAO global reported data in year 2006), the 633 transboundary species identified here accounted for 40.4 million tonnes of annual catches (i.e., 18% higher than previously estimated). Our results suggest that the contribution of transboundary species to global catch and fishing revenue might previously have been  underestimated due to an incomplete understanding of the transboundary nature of marine fished species. The 305 non-transboundary species (termed as ‘discrete’ species here, see Methods - Determining transboundary species trait) accounted for a much smaller proportion of total catch and revenue; 2.8 million tonnes and USD 4,282 million annually respectively, on average, between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

@@

At a regional level, I found that transboundary species are particularly economically important for Northern America (average country revenue = USD 4,680 $\pm$ 6,000 million) and Eastern Asia (USD 3,779 $\pm$ 3,093 million) (Figure \@ref(fig:ffvFigure1)b). The estimated per country revenues from transboundary species in these two regions is significantly different from other regions (one-way ANOVA; DF(16,165) = 5.081, p < 0.001, $\alpha$ = 0.05). China (USD 7,284 million) and the USA (USD 11,604 million) contribute 55% and 82% to the annual average revenue from 2005 to 2014  in Eastern Asia and Northern America, respectively. In addition to China and the USA, Russia (USD 7,379 million), Peru (USD 6,044 million) and Japan (USD 3,907 million) were among the top five countries with the most fishing revenue generated from transboundary species between 2005 and 2014 (Figure \@ref(fig:ffvFigure1)a). These five nations were responsible for 41% of the yearly global fisheries revenue from transboundary species.

@@

I determined the catch trend of each species within each EEZ, classifying them as increasing (Category A), constant (Category B) or decreasing (Category C) (Figures \@ref(fig:ffvFigure3); \@ref(fig:ffvFigureS2)). While previous studies have demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Our intention here is to examine where the catch trends of transboundary species differ from non-transboundary species between 2005 and 2014 relative to historic catch since 1951 (see Methods). I found significant differences for all catch trend categories for transboundary species (one-way ANOVA, Df(2,459) = 47.94, p < 0.001, $\alpha$ = 0.05; Tukey's post hoc test p $\leq$ 0.001), and no significant differences in catch trend categories for discrete species (one-way ANOVA, Df(2,106) = 1.885, p = 0.157; $\alpha$ = 0.05). I also found significant differences in catch trends when directly comparing transboundary to discrete species categories (MANOVA, Df(2,459) = 19.001, p < 0.001). Overall, transboundary species only targeted by one country are generally less likely to have a decreasing catch trend compared to instances where the shared species is fished by neighboring countries (Table \@ref(tab:ffvTableS2)).
